---
title: "Statistical Thinking Assignment 2"
author: "Xinyi Cui, Janice Hsin Hsu, Pranali Angne, Sahinya Akila"
date: "10/17/2021"
output: pdf_document
---

\newpage

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
# Loading Libraries

library(printr)
library(tidyverse)
library(tidymodels)
library(broom)
library(splines)
library(dagitty)
library(ggdag)
library(knitr)
library(gtsummary)
library(kableExtra)
```

# Task 1: Estimating Neonatal Mortality

## Introduction

One of this century’s global goals has been the reduction of childhood mortality across all countries. There has been enormous effort put into this goal at all levels from the united nations down to local interventions. The aim of this report is to produce a linear regression model to estimate the average neonatal mortality rate (NMR).

## Data

The source of the child mortality data is from the [UN Inter-agency Group for Child Mortality Estimation](https://childmortality.org/data).

It contains the following columns:

- `country_name`: Name of the country
- `year`: The year the data was measured
- `region`: The name of the continent the country is from
- `nmr`: The observed number of neonatal deaths per thousand live births (the neonatal mortality rate). This is measured either using a country’s vital registration system (births and deaths register) or using some sort of high-quality survey. 
- `u5mr`: The estimated under-five mortality rate
- `nmr_transformed`: log of the number of neonatal deaths per 1000 live births divided by the number of non-neonantal deaths per 1000 live births. 

$$log(\frac{nmr}{u5mr - nmr})$$
```{r}
# Reading the data
neonatal_mortality <- read_csv("neonatal_mortality.csv")

# Adding log ratio between neonatal mortality and non-neonatal mortality rate and time
nmr_data <- neonatal_mortality %>% 
  mutate(u5mr_log = log(u5mr), 
         transformed_nmr = log(nmr/(u5mr-nmr)),
         time = year - min(year))
```

## Task 1.1: Linear Regression

### Task 1.1.1: Explain the choice of variables in your model (you should not use country_name, if you use U5MR, you should use it on the log-scale!). In particular you should consider whether an interaction effect should be used.

```{r variable-choice, fig.width=20, fig.height=15}

# Splitting the data
nmr_split <- initial_split(nmr_data, strata = region)
nmr_train <- training(nmr_split)
nmr_test <- testing(nmr_split)

# Choosing the variables by trying different combinations of variables
fit1 <- lm(transformed_nmr ~ u5mr_log + region + time, nmr_train) %>% 
  tidy() %>% 
  mutate(model = "all")

fit2 <- lm(transformed_nmr ~ u5mr_log + region, nmr_train) %>% 
  tidy() %>% 
  mutate(model = "u5mr_region")

fit3 <- lm(transformed_nmr ~ u5mr_log + time, nmr_train) %>% 
  tidy() %>% 
  mutate(model = "u5mr_time")

fit4 <- lm(transformed_nmr ~ region + time, nmr_train) %>% 
  tidy() %>% 
  mutate(model = "region_time")

fit5 <- lm(transformed_nmr ~ region, nmr_train) %>% 
  tidy() %>% 
  mutate(model = "region")

fit6 <- lm(transformed_nmr ~ u5mr_log, nmr_train) %>% 
  tidy() %>% 
  mutate(model = "u5mr")

fit7 <- lm(transformed_nmr ~ time, nmr_train) %>% 
  tidy() %>% 
  mutate(model = "time")

# joining all the data frames and calculating the upper and lower values
full_model <- list(fit1, fit2, fit3, fit4, fit5, fit6, fit7) %>% 
  reduce(full_join) %>% 
  mutate(upper = estimate + 1.96 * std.error,
         lower = estimate - 1.96 * std.error)

full_model %>% ggplot(aes(estimate, y = model)) +
  geom_errorbar(aes(xmin = lower, xmax = upper, color = term)) +
  geom_vline(aes(xintercept = 0),
             linetype = "dashed",
             size = 1.2,
             color = "steel blue") +
  facet_wrap(.~term, scale = "free_x") +
  scale_y_discrete(limits = c("all", "u5mr_region", "u5mr_time", "region_time", "region", "u5mr", "time"))+ 
  theme(panel.background = element_rect(fill = "linen"), legend.position = "none", text = element_text(size=25))

```

In order to choose appropriate variables for the linear model, we have created seven models which includes different variables. Then we fitted seven different models into the error bar graph that represents the significance of each variable in the model.

From the error bar we can see that the time variable is considered to be significant among all models under the 95% confidence interval, as both the maximum and the minimum are different from 0. Similarly, variables such as region, and u5mr can also be considered as significant. However, while looking at the region_time model which represents the interaction effect between region and time, we can see that the shifts from the region model are only significant for High income and Latin American region; and therefore, the interaction effect should not be used here. 

For u5mr_region model, if we only consider u5mr or region, individually they are significant. When region is conditional on u5mr, the variable region turns from significant to insignificant, as it moves towards 0. Therefore, u5mr will affect region, but region have no effect on u5mr. We consider this is significant.

For u5mr_time model, as we mentioned before, u5mr itself is significant, while adding time, the significance level have not changed. However, when it is conditional on u5mr, time have been affected from positive values to negative significance values shown in graph t. Therefore, time u5mr will affect time. We also consider this is significant. 



```{r, fig.width=5, fig.height=5}

# understanding relationship between the variables
dagify(nmr ~ u5mr,
       nmr ~ time,
       nmr ~ Region,
       time ~ u5mr,
       Region ~ u5mr
       ) %>% 
  ggdag_adjust(node_size = 20) +
  theme_dag(legend.position = "none") 
  

# Fitting the model using the training data (with the interactive effect)
fit <- lm(transformed_nmr ~ u5mr_log + u5mr_log*region + u5mr_log*time, nmr_train)

summary(fit)
```
The dag graph illustrates The information about the relationship we get from the error bar graph. The neonatal mortality rate (nmr) is the dependent variable, from the dag graph we can see that both time and region is piped. We do not want to add them, cause they will end up worth less effect which removes the correlation between u5mr and nmr because we are adding bias through time and region. u5mr is a fork therefore we consider to include it. Moreover, from the error bar we see the indirect affect to u5mr and time, therefore, to consider the interaction effect, we include u5mr\*region and u5mr*time.

### Task 1.1.2: Assess your linear model and comment on its fit. This should be done a) for all data simultaneously; b)for data in each region; and c) for data in a maximum of 3 countries that should be chosen to highlight different aspects of the fit diagnostics.

```{r, fig.width=8}
augment_fit <- fit %>% 
  augment(data = nmr_train)
# part a
augment_fit %>% 
  ggplot() +
  geom_point(aes(x = u5mr_log, y= transformed_nmr), color = "#1b9ce3")+
  geom_line(aes(x = u5mr_log, y = .fitted), color = "#ffda85") +
  theme(panel.background = element_rect(fill = "#cae6da"), text = element_text(size=15)) +
  xlab("Under 5 mortality rate in Log scale") +
  ylab("Neonatal Mortality Rate in Log Scale")
  
```

The fitted model fits fairly well with all the data simultaneously, as we can see from the graph that the fitted model and the actual data mostly overlap with each other. Similarly, the fitted model works well with data in each region as well, especially for high income countries and Latin America and Caribbean. However, the fitted model does not work well with the countries Australia, India and Japan which we chose for highlight different aspects for the fit. The model only catches the trend in India. For more detailed fit diagnostics, we will assess the model based on the root mean square and the mean absolute error.

```{r, fig.width=20, fig.height=15}
# part b
augment_fit %>% 
  ggplot() +
  geom_point(aes(x = u5mr_log, y= transformed_nmr), color = "#7c9154")+
  geom_line(aes(x = u5mr_log, y = .fitted), color = "#ffda85")+
  facet_wrap(~region, scale = "free") +
  theme(panel.background = element_rect(fill = "#f0fcfc"), text = element_text(size=25)) +
  xlab("Under 5 mortality rate in Log scale") +
  ylab("Neonatal Mortality Rate in Log Scale")
```


```{r}
# part c
augment_fit %>%
  filter(country_name %in% c("India", "Japan", "Australia")) %>%
  ggplot(aes(x = u5mr_log, y = transformed_nmr)) +
  geom_point(color = "#126d7a") +
  geom_line(aes(x = u5mr_log, y = .fitted), color = "#ff8400") +
  facet_wrap(~country_name, scales = "free") +
  theme(panel.background = element_rect(fill = "#c3d3eb")) +
  xlab("Under 5 mortality rate in Log scale") +
  ylab("Fit")
```

We choose India, Australia and Japan to highlight different aspects of the fit. Australia and Japan both are developed countries, and India is a developing country. For Australia and Japan, they fit terribly for the model, especially Japan, non of the points are on the model. On contrast, India fits better than the other two countries.

### Task 1.1.3: Estimate the root mean square error and the mean absolute error on a test set. The test set should be produced using the argument strata = region.

```{r, fig.align='center'}

# Predict the test dataset
lm_pred <- tibble(pred = predict(fit, nmr_test))
lm_pred <- bind_cols(nmr_test, lm_pred)
lm_pred %>% 
  metrics(truth = transformed_nmr,
          estimate = pred) %>% 
  kbl(booktabs = T) %>%
  kable_styling(position = "center")
```

### Task 1.1.4: Produce a prediction, with prediction intervals, of the NMR on its natural scale (aka not on the log-scale) and plot these a) for all data simultaneously; b) for data in each region; and c) for data in a maximum of 3 countries that show different aspects of the fit.

```{r}
# Prediction for all data simultaneously
linear_prediction  <- as_tibble(predict(fit, nmr_test, interval = "prediction")) %>% 
  bind_cols(lm_pred)

linear_prediction  %>% 
  ggplot(aes(x = u5mr_log, y = transformed_nmr))+
  geom_point(color = "#287ad1")+
  geom_line(aes(y = lwr), color = "#d12828")+
  geom_line(aes(y = upr), color = "#d12828") +
  theme(panel.background = element_rect(fill = "#fff6ba")) +
  xlab("Under 5 mortality rate in Log scale") +
  ylab("Neonatal mortality rate")

#  Prediction for region
linear_prediction %>%
  ggplot(aes(x = u5mr_log, y = transformed_nmr))+
  geom_point()+
  geom_line(aes(y = lwr)) +
  geom_line(aes(y = upr))+
  facet_wrap(~ region, scale = "free")

#  Prediction for 3 Countires
 linear_prediction %>%
  filter(country_name %in% c("India", "Japan", "Australia")) %>% 
  ggplot(aes(x = u5mr_log, y = transformed_nmr))+
  geom_point()+
  geom_line(aes(y = lwr)) +
  geom_line(aes(y = upr))+
  facet_wrap(~ country_name, scale = "free")
```

## Task 1.2: Non-linear Regression

### Task 1.2.1: Explain your choice of model, using appropriate visualisations to support your choice. 

```{r}
#non-linear variable

#non-linear variable part1 choose variable


# find which variables have non-linear relationship
#logu5 and logr non linear relationship
ggplot(nmr_train)+
  geom_point(aes(x = u5mr_log, y = transformed_nmr), color = "#5905f5")+
  theme(panel.background = element_rect(fill = "#f0fcfc")) +
  xlab("Under 5 mortality rate in Log scale") +
  ylab("Neonatal Mortality Rate in Log Scale")

#time and logr, no non-linear relationship
ggplot(nmr_train)+
  geom_point(aes(x = time, y = transformed_nmr), color = "#5905f5")+
  theme(panel.background = element_rect(fill = "#f0fcfc")) +
  xlab("Years") +
  ylab("Neonatal Mortality Rate in Log Scale")

#region and logr, no non-linear
ggplot(nmr_train)+
  geom_point(aes(y = region, x = transformed_nmr), color = "#5905f5")+
  theme(panel.background = element_rect(fill = "#f0fcfc")) +
  ylab("Region") +
  xlab("Neonatal Mortality Rate in Log Scale")

# non-linear model
fit_non_linear <- lm(transformed_nmr ~ bs(u5mr_log, df= 15) + u5mr_log*region + u5mr_log*time, data = nmr_train)
  
```


### Task 1.2.2: Use cross-validation to select an appropriate number of basis functions for bs()

```{r}
#NOT DONE
```




### Task 1.2.3: For your final model, assess your linear model and comment on its fit. This should be done a) for all data simultaneously; b) for data in each region; and c) for data in a maximum of 3 countries that should be chosen to highlight different aspects of the fit diagnostics.
```{r}
# non-linear all
augment_non_linear <- fit_non_linear %>% 
  augment(data = nmr_train) 


augment_non_linear %>% 
  ggplot(aes(x = u5mr_log, y = transformed_nmr)) +
  geom_point()+
  geom_line(aes(x = u5mr_log, y = .fitted), color = "blue")

#non- linear region
augment_non_linear %>% 
  ggplot(aes(x = u5mr_log, y = transformed_nmr)) +
  geom_point()+
  geom_line(aes(x = u5mr_log, y = .fitted), color = "blue")+
  facet_wrap(~ region, scale = "free")

# 3 countries
augment_non_linear %>%
  filter(country_name %in% c("India", "Japan", "Australia")) %>%
  ggplot(aes(x = u5mr_log, y = transformed_nmr)) +
  geom_point(color = "#126d7a") +
  geom_line(aes(x = u5mr_log, y = .fitted), color = "#ff8400") +
  facet_wrap(~country_name, scales = "free") +
  theme(panel.background = element_rect(fill = "#c3d3eb")) +
  xlab("Under 5 mortality rate in Log scale") +
  ylab("Fit")

```



### Task 1.2.4: Estimate the root mean square error and the mean absolute error using the same test set as before.

```{r}
# Predict the test dataset
lm_pred <- tibble(pred = predict(fit_non_linear, nmr_test))
lm_pred <- bind_cols(nmr_test, lm_pred)
lm_pred %>% 
  metrics(truth = transformed_nmr,
          estimate = pred) %>% 
  kbl(booktabs = T) %>%
  kable_styling(position = "center")
```


### Task 1.2.5: Produce a prediction, with prediction intervals8, of the NMR on its natural scale (aka not on the log-scale) and plot these a) for all data simultaneously; b) for data in each region; and c) for data in a maximum of 3 countries that show different aspects of the fit.

```{r}
# Prediction for all data simultaneously
non_linear_prediction <- as_tibble(predict(fit_non_linear, nmr_test, interval = "prediction")) %>% 
  bind_cols(lm_pred)

non_linear_prediction %>% 
  ggplot(aes(x = u5mr_log, y = transformed_nmr))+
  geom_point(color = "#287ad1")+
  geom_line(aes(y = lwr), color = "#d12828")+
  geom_line(aes(y = upr), color = "#d12828") +
  theme(panel.background = element_rect(fill = "#fff6ba")) +
  xlab("Under 5 mortality rate in Log scale") +
  ylab("Neonatal mortality rate")

#  Prediction for region
non_linear_prediction %>%
  ggplot(aes(x = u5mr_log, y = transformed_nmr))+
  geom_point()+
  geom_line(aes(y = lwr)) +
  geom_line(aes(y = upr))+
  facet_wrap(~ region, scale = "free")

#  Prediction for 3 Countires
 non_linear_prediction %>%
  filter(country_name %in% c("India", "Japan", "Australia")) %>% 
  ggplot(aes(x = u5mr_log, y = transformed_nmr))+
  geom_point()+
  geom_line(aes(y = lwr)) +
  geom_line(aes(y = upr))+
  facet_wrap(~ country_name, scale = "free")
```




