---
title: "ETC5242Assignment"
author: "Sahinya Akila"
date: "9/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(survival)
library(survminer)
library(kableExtra)
library(knitr)
library(ggplot2)
## Remove the line break in the file name!
churn_dat <- read_csv("https://raw.githubusercontent.com/square/pysurvival/master/pysurvival/datasets/churn.csv")
churn_dat <- churn_dat %>% filter(months_active > 0) %>% select(c(company_size, months_active, churned))
```

```{r}
km_model <- function(time, event){
  dataset <- data_frame(time, event)
  
  km_data <- dataset %>% 
    group_by(time, event) %>% 
    summarise(died = n()) %>% 
    ungroup() %>% 
    mutate(risk = nrow(dataset) - accumulate(died, `+`) + died) %>% 
    filter(event == 1) %>% 
    mutate(probability = 1 - died/risk,
           survival = accumulate(probability, `*`))
  return(km_data %>% select(time, survival))
}

km_survive <- km_model(churn_dat$months_active, churn_dat$churned) 

km_survive %>% 
  ggplot(aes(time, survival)) +
  geom_step()
```


```{r}
company_km_model <- data.frame(time = double(), survival = double(), company_size = character())
for(size in unique(churn_dat$company_size)){
  filtered <- churn_dat %>% filter(company_size == size)
  final_model <- km_model(filtered$months_active, filtered$churned) %>% mutate(company_size = size)
  company_km_model <- rbind(company_km_model, final_model)
}

company_km_model %>% 
  ggplot(aes(time, survival)) +
  geom_step() +
  facet_wrap(~company_size)
```


# Question 2


## Compute the Kaplan-Meir curve and use this to estimate the median churn time.

```{r}
fit <- survfit(Surv(months_active, churned) ~ company_size,
                  data = churn_dat)

g2 <- ggsurvplot(fit,conf.int = FALSE)

g2$plot +
  labs(x = "Months Active",
       title = "Kaplan-Meir curve for each company size") +
  theme(legend.position="bottom") 
```


```{r}
median_function <- function(fit){
  index <- which.min(log(abs(fit$surv - 0.5)))
  median <- fit$time[index]
  return(median)
}
```


```{r}

for (size in unique(churn_dat$company_size)){
  temp_data <- churn_dat %>% filter(company_size == size)
  name <- size
  assign(name, survfit(Surv(months_active, churned) ~ company_size, data = temp_data))
}

```

```{r}
company_median <- data_frame(company_size = unique(churn_dat$company_size), median = c(NA, NA, NA, NA, NA))

for (i in 1:length(company_median$company_size)){
  company_median$median[i] <- median_function(get(company_median$company_size[i]))
}

company_median %>%
  knitr::kable(
  caption = "Medians for different company sizes") %>%
  kable_styling(c("hover", "striped"))
```

The table above demonstrates the median churn time estimated for different company size.
- Company size of 1-10 have the highest estimated median of 7 months.
- Company size of 100-250 have the lowest estimated median of 4 months.
- The rest of the company sizes have the same estimated median of 5 months.

## Use a non-parametric bootstrap to construct 90% confidence intervals for the median of each company size


```{r}
bootstrapmedian <- function(df_median, df){
  bootstrap <- tibble(experiment = rep(1:1000, each = nrow(df)),
                        ind = sample(1:nrow(df), size = nrow(df)*1000, replace = TRUE),
                        timestar = df$months_active[ind],
                        churnstar = df$churned[ind])
  
  bias <- bootstrap %>%
    group_by(experiment) %>%
    summarise(delta = median_function(df_median) - median_function(survfit(Surv(timestar, churnstar) ~ experiment)))

  ci <- median_function(df_median) + quantile(bias$delta, c(0.05, 0.95))
  
  return(ci)
}
```

```{r}
company_median_ci <- data_frame(company_size = unique(churn_dat$company_size), median = c(NA, NA, NA, NA, NA), lcl = c(NA, NA, NA, NA, NA), ucl = c(NA, NA, NA, NA, NA))

for (i in 1:length(company_median_ci$company_size)){
  ci <- bootstrapmedian(get(company_median_ci$company_size[i]), churn_dat %>% filter(company_size == company_median_ci$company_size[i]))
  company_median_ci$median[i] <- median_function(get(company_median_ci$company_size[i]))
           company_median_ci$lcl[i] <- ci[1]
           company_median_ci$ucl[i] = ci[2]
}
company_median_ci %>%
  knitr::kable(
  caption = "estimated mean under 90% CI") %>%
  kable_styling(c("hover", "striped"))
```
## Make a plot that shows that estimate of the median and the corresponding confidence interval on the same axes

```{r}
ggplot(company_median_ci,
       aes(x = company_size,
           y = median,
           colour = company_size)) + 
  geom_errorbar(aes(ymax = ucl, ymin = lcl)) +
  geom_point() +
  theme_bw() +
  labs(x = "Company Size",
       y = "Estimated Median Churn Time",
       title = "estimated mean under 90% CI")
```

