---
title: "ETC5242Assignment"
author: "Sahinya Akila"
date: "9/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(survival)
library(survminer)
## Remove the line break in the file name!
churn_dat <- read_csv("https://raw.githubusercontent.com/square/pysurvival/master/pysurvival/datasets/churn.csv")
churn_dat <- churn_dat %>% filter(months_active > 0) %>% select(c(months_active, churned, company_size))
```

```{r}
km_model <- function(time, event){
  dataset <- data_frame(time, event)
  
  km_data <- dataset %>% 
    group_by(time, event) %>% 
    summarise(died = n()) %>% 
    ungroup() %>% 
    mutate(risk = nrow(dataset) - accumulate(died, `+`) + died) %>% 
    filter(event == 1) %>% 
    mutate(probability = 1 - died/risk,
           survival = accumulate(probability, `*`))
  return(km_data %>% select(time, survival))
}

km_survive <- km_model(churn_dat$months_active, churn_dat$churned) 

km_survive %>% 
  ggplot(aes(time, survival)) +
  geom_step()
```


```{r}
company_km_model <- data.frame(time = double(), survival = double(), company_size = character())
for(size in unique(churn_dat$company_size)){
  filtered <- churn_dat %>% filter(company_size == size)
  final_model <- km_model(filtered$months_active, filtered$churned) %>% mutate(company_size = size)
  company_km_model <- rbind(company_km_model, final_model)
}

company_km_model %>% 
  ggplot(aes(time, survival)) +
  geom_step() +
  facet_wrap(~company_size)
```


# Q2


- Compute the Kaplan-Meir curve and use this to estimate the median churn time.

```{r}
fit <- survfit(Surv(months_active, churned) ~ company_size,
                  data = churn_dat)

g2 <- ggsurvplot(fit,conf.int = FALSE)

g2$plot +
  labs(x = "Months Active",
       title = "Kaplan-Meir curve for each company size") +
  theme(legend.position="bottom") 
```

We calculate the estimated median churn time based where fit$surv is closest to 0.5.

```{r}
median_function <- function(x){
  index <- which.min(log(abs(fit$surv - 0.5)))
  median <- fit$time[index]
  return(median)
}
```

```{r}

for (size in unique(churn_dat$company_size)){
  temp_data <- churn_dat %>% filter(company_size == size)
  name <- paste(size, "fit", sep = "_")
  assign(name, survfit(Surv(months_active, churned) ~ company_size, data = temp_data))
}

```

```{r}
company_median <- data_frame(company_size = unique(churn_dat$company_size))

for (i in 1:length(company_median)){
  company_median <- company_median %>% mutate(median = median_function(paste(company_median[i], "fit", sep="_")))
}
```

- Use a non-parametric bootstrap to construct 90% confidence intervals for the median of each company
size

A bootmedian function is created to conduct non-parametric bootstrap and calculate 90% confidence interval for each company size, 1000 bootstrap samples are drew for each company size.

```{r}
bootstrapmedian <- function(df_median, df){
  bootstrap <- tibble(experiment = rep(1:1000, each = nrow(df)),
                        ind = sample(1:nrow(df), size = nrow(df)*1000, replace = TRUE),
                        timestar = df$months_active[ind],
                        churnstar = df$churned[ind])
  
  bias <- bootstrap %>%
    group_by(experiment) %>%
    summarise(delta = near_median(df_median) - near_median(survfit(Surv(timestar, churnstar) ~ experiment)))

  ci <- near_median(df_median) + quantile(bias$delta, c(0.05, 0.95))
  
  return(ci)
}
```

```{r}
company_median_ci <- data_frame(company_size = unique(churn_dat$company_size))
for (i in 1:length(company_median_ci)){
  ci <- bootstrapmedian(paste(company_median[i], "fit", sep="_"), churn_dat %>% filter(company_size == company_median_ci[i,]))
  company_median_ci <- company_median_ci %>% 
    mutate(median = median_function(paste(company_median[i], "fit", sep="_")),
           lcl = ci[1], ucl = ci[2])
}
```

```{r}
q2_medianci %>% 
  kable(col.names = c("Company Size", "Estimated Median Churn Time",
                      "Lower Confidence Interval", "Upper Confidence Interval"),
      align = "c",
      caption = "90% Confidence interval for estimated median - non-parametric bootstrap") %>% 
  kable_styling(bootstrap_options = c("striped","hover"))
```

- Make a plot that shows that estimate of the median and the corresponding confidence interval on the
same axes

```{r}
ggplot(q2_medianci,
       aes(x = reorder(company_size, median),
           y = median,
           colour = company_size)) + 
  geom_point(size = 6) +
  geom_errorbar(aes(ymax = ucl, ymin = lcl)) +
  theme_bw() +
  labs(x = "Company Size",
       y = "Estimated Median Churn Time",
       title = "90% Confidence interval for estimated median - non-parametric bootstrap") +
  coord_flip()
  
```

- Write some sentences describing how the median churn time changes across company sizes.

The plots and table above demonstrate that the median churn time is the lowest for company size 100-250, which is only 4 months, and company size of 1-10 has the highest estimated median churn time of 7 months. As for company size of 10-50, 50-100 and self-employed, the estimated median churn times are the same, which is 5 months. 


